---
title: "2023 Business Counts of Local Units for South Yorkshire, Greater Manchester, West Yorkshire and Merseyside (Liverpool City Region) broken down by broad industrial group"
author: "Dan Olner"
date: "2023-10-13"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
#Don't think DPI makes a difference... oh yes it does, ace (not in Word though?)
knitr::opts_chunk$set(echo = F, warning = F, error = F, message = F, comment=NA, dpi=300, fig.width = 9)

library(tidyverse)
library(cowplot)
```

```{r}
lu23 <- read_csv('../local/data/BusinessCountsByNUTS2/manualdownload_NUTS2_localunits_2023_readyforimport.csv') %>% 
  pivot_longer(cols = `UKC1 Tees Valley and Durham`:`UKN0 Northern Ireland`, names_to = 'GEOGRAPHY_NAME', values_to = 'COUNT') %>% 
  mutate(
    GEOGRAPHY_NAME = substring(GEOGRAPHY_NAME, 6)#Drop codes
  ) 
  
#Industries are in the correct order in that load. Get factor order from there.
industryorder <- unique(lu23$Industry)[1:length(unique(lu23$Industry))-1]


#OK, so just getting a picture for this year.
#We want proportions of total business counts, so...

#But let's start with totals for ALL industries to get overview
lu.all <- lu23 %>% 
  filter(Sizeband!='All') %>% #otherwise, double counting!
  group_by(Sizeband,GEOGRAPHY_NAME) %>% 
  summarise(COUNT = sum(COUNT)) %>% 
  group_by(GEOGRAPHY_NAME) %>% 
  mutate(
    REGION_TOTAL = sum(COUNT),
    REGION_PERCENT = (COUNT/REGION_TOTAL) * 100
    ) %>% 
  mutate(Sizeband = factor(Sizeband, levels = c("Micro (0 to 9)","Small (10 to 49)","Medium-sized (50 to 249)","Large (250+)")))

#sanity check on region percents
# lu.all %>% group_by(GEOGRAPHY_NAME) %>% summarise(sum(REGION_PERCENT))

#Merseyside is liverpool city region, checked, matches
#See map at https://en.wikipedia.org/wiki/Liverpool_City_Region

#List of comparators
places = unique(lu.all$GEOGRAPHY_NAME[grepl('Merseyside|South York|Manchester|West York', lu.all$GEOGRAPHY_NAME)])

#Repeat for the broad industrial groups
lu.ind <- lu23 %>% 
  filter(Sizeband!='All', Industry!='Column Total') %>% #otherwise, double counting!
  group_by(Sizeband,GEOGRAPHY_NAME,Industry) %>% 
  summarise(COUNT = sum(COUNT)) %>% 
  group_by(GEOGRAPHY_NAME) %>% 
  mutate(
    REGION_TOTAL = sum(COUNT),
    REGION_PERCENT = (COUNT/REGION_TOTAL) * 100
  ) %>% 
  mutate(
    Sizeband = factor(Sizeband, levels = c("Micro (0 to 9)","Small (10 to 49)","Medium-sized (50 to 249)","Large (250+)")),
    Industry = factor(Industry, levels = industryorder)
    )
```


# Total business counts for each region

* The four regions are broadly similar in percent of business counts per size band.
* Micro firms for all sectors make up by far the largest proportion of local units.
* Manchester slightly ahead for micro, LCR and SY for small and medium-size firms (but suspect error bars might wipe that difference out)

```{r}

#OK, so with totals, basic comparison
ggplot(
  lu.all %>% filter(GEOGRAPHY_NAME %in% places), 
  aes(x = Sizeband, y = REGION_PERCENT, fill = GEOGRAPHY_NAME)
) +
  geom_bar(stat='identity', position = 'dodge') +
  xlab("Business size band") +
  ylab("Percent of region's total business count") +
  scale_fill_brewer(palette = "Paired", direction = 1) +
  theme(legend.title=element_blank())



```



# Each size band with its own scale

**Zooming in to each size band to make the (not very large) differences a bit more clear. The business count numbers are rounded to 5, and removed if not enough values - particularly tricky for 250+. See below to see 250+ counts.**

```{r}
#Repeat and separate out to compare scale of each sizeband
ggplot(
  lu.all %>% filter(GEOGRAPHY_NAME %in% places), 
  aes(x = Sizeband, y = REGION_PERCENT, fill = GEOGRAPHY_NAME)
) +
  geom_bar(stat='identity', position = 'dodge') +
  facet_wrap(~Sizeband, scales = 'free') +
  xlab("Business size band") +
  ylab("Percent of region's total business count") +
  scale_fill_brewer(palette = "Paired", direction = 1) +
  theme(legend.title=element_blank())

```

# Next slide: Business counts by industry, overview

**The next slide has business counts by industry for each region, with the x axis fixed so all sectors and regions can be compared.**

* Certain sectors much more micro-firm heavy
* SY/WY a little ahead in manufacturing micro and small firms (see below for more on that)
* Proportionally many more SY micro-firms for transport/storage
* Fewer SY micro firms in 13, 14 and 17



***   

```{r fig.height = 9, fig.width = 17}
#Changing dimensions helps, DPI does not
# ```{r dpi = 600}
ggplot(
  lu.ind %>% filter(GEOGRAPHY_NAME %in% places), 
  aes(x = Sizeband, y = REGION_PERCENT, fill = GEOGRAPHY_NAME)
) +
  geom_bar(stat='identity', position = 'dodge') +
  facet_wrap(~Industry) +
  coord_flip() +
  xlab("Business size band") +
  ylab("Percent of region's total business count") +
  scale_fill_brewer(palette = "Paired", direction = 1) +
  theme(legend.title=element_blank())
  

```


# Next slide: Business counts by industry, overview - free percent axis

**Same again, but with the percent axis free, so e.g. we can see for public admin and defence (15) what the mix of business sizes is a little more clearly.**


***   


```{r fig.height = 9, fig.width = 17}
ggplot(
  lu.ind %>% filter(GEOGRAPHY_NAME %in% places), 
  aes(x = Sizeband, y = REGION_PERCENT, fill = GEOGRAPHY_NAME)
) +
  geom_bar(stat='identity', position = 'dodge') +
  facet_wrap(~Industry, scales = 'free_x') +
  coord_flip() +
  xlab("Business size band") +
  ylab("Percent of region's total business count") +
  scale_fill_brewer(palette = "Paired", direction = 1) +
  theme(legend.title=element_blank())

```



# Next slide: stacked bars showing composition of each industry size

**Public sectors have far higher numbers of large organisations, predictably. Stand-outs are South Yorkshire's higher proportion of micro-firms / lower for 250+ firms in public admin & defence.**


***   




```{r fig.height = 9, fig.width = 17}
ggplot(
  lu.ind %>% filter(GEOGRAPHY_NAME %in% places), 
  aes(x = GEOGRAPHY_NAME, y = REGION_PERCENT, fill = Sizeband)
) +
  geom_bar(stat='identity', position = 'fill') +
  facet_wrap(~Industry) +
  coord_flip() +
  xlab("Region") +
  ylab("Percent of region's total business count") +
  scale_fill_brewer(palette = "Paired", direction = 1) +
  theme(legend.title=element_blank())

```


# Next slide: Micro (0-9) units broken down by industry

**Looking just at micro-size local units (0 to 9 people), a few South Yorkshire stand-outs:**

* Larger % of construction micro-firm units than the comparators
* Much larger % for transport and storage
* GM a way ahead in finance/insurance at this scale
* Professional, scientific & technical - lower for SY.
* (Public sector units generally much lower micro-firm numbers.)

***   



```{r fig.height = 9, fig.width = 17}
ggplot(
  lu.ind %>% filter(GEOGRAPHY_NAME %in% places, Sizeband == 'Micro (0 to 9)'), 
  aes(x = Sizeband, y = REGION_PERCENT, fill = GEOGRAPHY_NAME)
) +
  geom_bar(stat='identity', position = 'dodge') +
  facet_wrap(~Industry, scales = 'free_x')+
  scale_fill_brewer(palette = "Paired", direction = 1) +
  xlab("Business size band") +
  ylab("Percent of region's total business count") +
  theme(legend.title=element_blank(), plot.title = element_text(face = 'bold'), axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ggtitle("Micro (0-9)")
```


# Next slide: Small (10 to 49) units broken down by industry

* More in SY for manufacturing and motor trades
* SY a stand out as larger in public admin and defence at this scale

***   




```{r fig.height = 9, fig.width = 17}
ggplot(
  lu.ind %>% filter(GEOGRAPHY_NAME %in% places, Sizeband == 'Small (10 to 49)'), 
  aes(x = Sizeband, y = REGION_PERCENT, fill = GEOGRAPHY_NAME)
) +
  geom_bar(stat='identity', position = 'dodge') +
  facet_wrap(~Industry, scales = 'free')+
  scale_fill_brewer(palette = "Paired", direction = 1)+
  xlab("Business size band") +
  ylab("Percent of region's total business count") +
  theme(legend.title=element_blank(), plot.title = element_text(face = 'bold'), axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ggtitle("Small (10 to 49)")

```


# Next slide: Medium-sized (50 to 249) units broken down by industry

* SY and WY both much higher % in manufacturing here, same for motor trades.
* Finance: SY lower than elsewhere for medium size firms



***   



```{r fig.height = 9, fig.width = 17}
ggplot(
  lu.ind %>% filter(GEOGRAPHY_NAME %in% places, Sizeband == 'Medium-sized (50 to 249)'), 
  aes(x = Sizeband, y = REGION_PERCENT, fill = GEOGRAPHY_NAME)
) +
  geom_bar(stat='identity', position = 'dodge') +
  facet_wrap(~Industry, scales = 'free')+
  scale_fill_brewer(palette = "Paired", direction = 1)+
  xlab("Business size band") +
  ylab("Percent of region's total business count") +
  theme(legend.title=element_blank(), plot.title = element_text(face = 'bold'), axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ggtitle("Medium-sized (50 to 249)")

```


# Next slide: Actual counts for 250+ local units, by industry

**Because there are far fewer 250+ size sites, the applied rounding to 5 affects them a lot more, meaning percents are dubious. Looking at the raw counts is a useful alternative (though caution needed given they aren't adjusted for region size).**

**A South Yorkshire stand out here is the very small number of large professional/scientific/technical local units, matching lower micro numbers.**

***   


```{r fig.height = 9, fig.width = 17}
ggplot(
  lu.ind %>% filter(GEOGRAPHY_NAME %in% places, Sizeband == 'Large (250+)'), 
  aes(x = Sizeband, y = COUNT, fill = GEOGRAPHY_NAME)
) +
  geom_bar(stat='identity', position = 'dodge') +
  facet_wrap(~Industry) +
  # facet_wrap(~Industry, scales = 'free') +
  xlab("") +
  ylab("Total business count for 250+ firms") +
  scale_fill_brewer(palette = "Paired", direction = 1) +
  theme(legend.title=element_blank(), plot.title = element_text(face = 'bold'), axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ggtitle("Counts of 250+ size local units")
```



# Next slide: Picking out some of the stand out sectors from above, putting together

* GM far ahead in finance in micro and medium size, SY behind across the board
* Professional/scientific: SY especially low in large firms (but note issue of counts rounding in large firm numbers)
* SY Public admin / defence numbers higher in micro/small - what kinds of firms are those?
* Manufacturing East / West split, SY and WY ahead in all firm sizes
* Construction more mixed, with SY ahead for both ends - micro and large, far ahead for large
* SY ahead for most transport/storage firm scales, LCR in medium-sized though.

***   



```{r fig.height = 9, fig.width = 17}
bespoke <- unique(lu.ind$Industry[grepl(x = lu.ind$Industry, pattern = 'manuf|construction|transport|financ|scient|defence', ignore.case = T)])

plotindustry <- function(industryname){
  
  p <- ggplot(
    lu.ind %>% filter(GEOGRAPHY_NAME %in% places, Industry == industryname), 
    aes(x = Sizeband, y = REGION_PERCENT, fill = GEOGRAPHY_NAME)
  ) +
    geom_bar(stat='identity', position = 'dodge') +
    facet_wrap(~Sizeband, scales = 'free')+
    scale_fill_brewer(palette = "Paired", direction = 1) +
    xlab("Business size band") +
    ylab("Percent of region's total business count") +
    scale_fill_brewer(palette = "Paired", direction = 1) +
    theme(legend.title=element_blank(), plot.title = element_text(face = 'bold')) +
    ggtitle(industryname)
  
}

plots <- map(bespoke, plotindustry)

cp <- plot_grid(plotlist = plots, nrow = 2)
cp

```


