---
title: "Looking at the 20 broad sector 'SIC sections' and picking out significant growth sectors"
author: "Dan Olner"
date: "2023-11-27"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
#Don't think DPI makes a difference... oh yes it does, ace (not in Word though?)
knitr::opts_chunk$set(echo = F, warning = F, error = F, message = F, comment=NA, fig.width = 12, fig.height = 12)

library(tidyverse)
library(cowplot)
library(zoo)
library(sf)
library(tmap)
library(flextable)
source('../functions/misc_functions.R')
options(scipen = 99)#Avoids scientific notation
```

```{r loaddata}
itl2.cvs <- readRDS('../data/UKchainedvolume_itl2_SIC_sections.rds')

itl2.jobs <- readRDS('../data/itl2_BRES_jobs_SIC_sections.rds')

#Mark South Yorkshire
itl2.cvs$ITL_region_name[itl2.cvs$ITL_region_name == 'South Yorkshire'] <- 'SOUTH YORKSHIRE <<<'
itl2.jobs$GEOGRAPHY_NAME[itl2.jobs$GEOGRAPHY_NAME == 'South Yorkshire'] <- 'SOUTH YORKSHIRE <<<'

#Right join to match the fewer available years in the BRES data
itl2.gvaperjob <- itl2.cvs %>% 
  rename(gva = value) %>% 
  right_join(
    itl2.jobs %>% select(-SIC_SECTION_NAME) %>% rename(jobcount = COUNT),
    by = c('year' = 'DATE','ITL_region_name' = 'GEOGRAPHY_NAME','SIC07_code' = 'SIC_SECTION_CODE')
  ) %>% 
  mutate(gvaperjob = (gva/jobcount) * 1000000)


slopes.log1521 <- get_slope_and_se_safely(data = itl2.cvs %>% filter(year %in% 2015:2021), ITL_region_name,SIC07_description, y = log(value), x = year)
slopes.log1319 <- get_slope_and_se_safely(data = itl2.cvs %>% filter(year %in% 2013:2019), ITL_region_name,SIC07_description, y = log(value), x = year)

```


# AIMS FOR THESE SLIDES

-- Look at jobs and GVA for South Yorkshire, comparing to the rest of UK.

-- By comparing slopes of change over time, pull out growth that statistically stands out from other sectors and places. (These sectors will also be ones whose linear growth trends would stand out projected forward in time.)

-- GVA here is 'chained volume' i.e. it's inflation-adjusted and reflects actual per-sector output. Job counts are from the Business Register and Employment Survey (BRES)

----

# HEADLINES

-- Information and Communication (ICT) is the stand out sector, given this data. Its growth stands out statistically through both a jobs and GVA lens.

-- Manufacturing is strong, but it's less statistically distinct from other places in the UK.

-- No other broad SIC section is strongly statistically separable from others.


----

# OVERVIEW OF THE JOB COUNT VS GVA DATA JUST FOR SOUTH YORKSHIRE

The next two slides show the scale of job counts and GVA values for South Yorkshire, comparing 2015 to 2021 and showing how the composition of 'GVA per job' has changed in very different ways for different sectors.

After this, we'll look more closely at change trends and try to separate out what's statistically interesting.

----

## Job counts vs GVA for South Yorkshire

The GVA per worker data is volatile because of the uncertainty in the BRES data. But there are some useful insights from plotting GVA vs job counts directly and seeing how they have changed between the first and last year in the data (2015 and 2021). 

GVA per full time job figures (the x axis divided by the y axis) for the start and end year are included in the labels to show how it's changed.

Sector breakdown in the next slide. Note, none of the values have error rates, but there are still relevant patterns...


```{r}
p <- twod_generictimeplot(
  df = itl2.gvaperjob %>% filter(ITL_region_name == 'SOUTH YORKSHIRE <<<', SIC07_description!='Real estate activities') %>% mutate(`gva/job` = gvaperjob/1000), 
  category_var = SIC07_description,
  x_var = gva,
  y_var = jobcount,
  timevar = year,
  label_var = `gva/job`,
  start_time = 2015,
  end_time = 2021
)

p + theme(aspect.ratio=1)

```

----

## Job counts vs GVA for South Yorkshire

-- Information & communication (ICT), bottom left going 'north east', is one of the few sectors here with both positive GVA and jobs growth, and GVA per worker has increased too.

-- Manufacturing GVA grew, job counts between 2015 and 21 hardly changed

-- Human health saw a GVA drop alongside jobs growth: the drop in GVA per worker is doubly stark.

-- Transportation/storage has seen a similar (though not as steep) jobs increase coupled with GVA drop

-- Admin/support services is an example of GVA per worker increase through apparent job drops while maintaining the same output (opposite of manufacturing)


```{r}
p <- twod_generictimeplot(
  df = itl2.gvaperjob %>% filter(ITL_region_name == 'SOUTH YORKSHIRE <<<', SIC07_description!='Real estate activities') %>% mutate(`gva/job` = gvaperjob/1000), 
  category_var = SIC07_description,
  x_var = gva,
  y_var = jobcount,
  timevar = year,
  label_var = `gva/job`,
  start_time = 2015,
  end_time = 2021
)

p + theme(aspect.ratio=1)

```


----

# GVA trends

The next slides look at trends in GVA (using linear regression lines) and try to separate out different slopes to identify true growth sectors. Very few can be strongly statistically separated.

----

## Example of separating trend lines

This plot shows construction and manufacturing chained volume (inflation-adjusted) GVA for South Yorkshire, with 95% confidence intervals shown in grey. The two trends are clearly statistically separable - the range of the trend slopes within the grey bounds differs; additionally, manufacturing has a positive slope and mining/quarrying is negative.

In the grid plot for South Yorkshire on the following slide, each square is a pair of slopes like this. As the slide explains, it indicates where pairs of slopes are statistically separarable, to help see which sectors and places are most likely clearly growing/shrinking more than others.

```{r}
#Sample plot comparing two slopes to show sig diff
#manuf vs mining quarrying
sampleplot <- itl2.cvs %>%
  filter(
    ITL_region_name == 'SOUTH YORKSHIRE <<<',
    # SIC07_description %in% c('Manufacturing','Mining and quarrying')
    SIC07_description %in% c('Manufacturing','Construction')
  )

#default level is 0.95
ggplot(sampleplot, aes(x=year,y=value,colour=SIC07_description )) +
  geom_line(size=2) +
  # scale_y_log10() +
  geom_smooth(method='lm') 
```

----

## South Yorkshire trend matrix for 20 SIC sections' trendlines 2015 to 2021

This matrix takes every pair of trendlines in South Yorkshire and compares them. How to read:

-- Green and red squares show the DIFFERENCE between slope lines (so green may not mean one slope is positive, just that the difference is positive) 

-- If a pair's square has a black outline, the difference is statistically separable (CI 95%)

-- Colours of the axis names indicate whether that sector's slope was positive (green vs red) and whether significant (95% CI, in bolder green or red).

-- The matrix is mirrored along the diagonal axis so e.g. ICT's green horizontal line is mirrored as red vertically (showing the same thing - ICT higher slopes than other sectors, other sectors lower than ICT).

Information & communication is the clear standout, with manufacturing and the power and water sectors separable from others the most. But note, there's a COVID effect here... see next slide of the same plot for 2013-2019. (Treat real estate activities cautiously - it includes imputed rental.)


```{r}
p <- slopeDiffGrid(slope_df = slopes.log1521, confidence_interval = 95, column_to_grid = SIC07_description, column_to_filter = ITL_region_name, filterval = 'SOUTH YORKSHIRE <<<')

p + coord_fixed()

```

----

## South Yorkshire trend matrix for 20 SIC sections' trendlines 2013 to 2019

Looking at the 7 years prior to COVID, most of the separable sectors disappear - Information/communication remain, however.

This is showing how some sectors were more resilient through COVID than others, i.e. in the previous plot, manufacturing slopes are more separable because it did better compared to others through COVID. Slides below look at this in more detail.


```{r}
# slopes.log <- get_slope_and_se_safely(data = itl2.cvs, ITL_region_name,SIC07_description, y = log(value), x = year)
slopes.log <- get_slope_and_se_safely(data = itl2.cvs %>% filter(year %in% 2013:2019), ITL_region_name,SIC07_description, y = log(value), x = year)

p <- slopeDiffGrid(slope_df = slopes.log1319, confidence_interval = 95, column_to_grid = SIC07_description, column_to_filter = ITL_region_name, filterval = 'SOUTH YORKSHIRE <<<')

p + coord_fixed()

```



----

## Information & Communication time series: 10 HIGHEST growth places 2015-21

Looking just at Information/communication, this plot shows the top 10 growth places (with South Yorkshire overlaid) using slopes between 2015-21 - though displaying the whole data range to show ICT has been constantly growing (as the next lowest-growth slide also shows).

But even given that ICT growth has been ubiquitous, South Yorkshire's output stands out. It jumped from 2018 and was less affected by COVID than other places. This has put its ICT growth statistically ahead of most (all?) other places.

Note: the raw GVA amounts for each place have NOT been adjusted e.g. on a per capita or worker basis, so direct scale comparisons cannot be made - compare change over time only.

```{r, fig.width = 10, fig.height = 8}
growing = T
place = 'SOUTH YORKSHIRE <<<'
sector = 'Information and communication'

timeplot <- itl2.cvs %>% 
      filter(SIC07_description == sector) 
    
#Use zoo's rollapply function to get a moving average
timeplot <- timeplot %>% 
  group_by(ITL_region_name) %>% 
  arrange(year) %>% 
  mutate(
    movingav = rollapply(value,3,mean,align='right',fill=NA)
  )

#Use largest positive log slopes to filter
if(growing){
  place_selection <- slopes.log1521 %>% filter(SIC07_description == sector) %>%
    arrange(-slope)
  sy_position <<- which(place_selection$ITL_region_name == 'SOUTH YORKSHIRE <<<')
} else {
  place_selection <- slopes.log1521 %>% filter(SIC07_description == sector) %>%
    arrange(slope)
    sy_position <<- length(place_selection$ITL_region_name) + 1 -which(place_selection$ITL_region_name == 'SOUTH YORKSHIRE <<<')
  }

#Keep only the top ten places and order them
timeplot <- timeplot %>% 
  mutate(ITL_region_name = factor(ITL_region_name, ordered = T, levels = place_selection$ITL_region_name)) %>%
  filter(ITL_region_name %in% c(place_selection$ITL_region_name[1:10],place))#Make sure chosen place added

#Mark the ITL of interest so it can be clearer in the plot
timeplot <- timeplot %>%
  mutate(
    ITL2ofinterest = ifelse(ITL_region_name == place, 'ITL of interest','other'),
  )

ggplot(timeplot %>% 
         rename(`ITL region` = ITL_region_name) %>% 
         filter(!is.na(movingav)),#remove NAs from dates so the x axis doesn't show them
       aes(x = year, y = value, colour = `ITL region`, size = ITL2ofinterest, linetype = ITL2ofinterest, group = `ITL region`)) +
       # aes(x = year, y = movingav, colour = `ITL region`, size = ITL2ofinterest, linetype = ITL2ofinterest, group = `ITL region`)) +
  geom_point() +
  geom_line() +
  scale_y_log10() +
  scale_size_manual(values = c(2.5,1)) +
  scale_color_brewer(palette = 'Paired', direction = 1) +
  ylab('Regional GVA chained volume (log, millions)') +
  guides(size = "none", linetype = "none") +
  ggtitle(
    paste0(ifelse(growing,'10 HIGHEST GROWTH PLACES','10 LOWEST GROWTH PLACES'),'\n',sector,'\n', place, ' highlighted (',scales::ordinal(sy_position),' most positive slope)')
  ) +
  theme(plot.title = element_text(face = 'bold'))
```

----

## Information & Communication time series: 10 LOWEST growth places

These 10 places had the lowest growth slopes for ICT in the UK - and are still all growing. This is the national context for this sector.

Note how clearly the effect of COVID is visible from 2019 onwards for many of these places (including in the previous slide).

```{r, fig.width = 10, fig.height = 8}
growing = F
place = 'SOUTH YORKSHIRE <<<'
sector = 'Information and communication'

timeplot <- itl2.cvs %>% 
      filter(SIC07_description == sector) 
    
#Use zoo's rollapply function to get a moving average
timeplot <- timeplot %>% 
  group_by(ITL_region_name) %>% 
  arrange(year) %>% 
  mutate(
    movingav = rollapply(value,3,mean,align='right',fill=NA)
  )

#Use largest positive log slopes to filter
if(growing){
  place_selection <- slopes.log1521 %>% filter(SIC07_description == sector) %>%
    arrange(-slope)
  sy_position <<- which(place_selection$ITL_region_name == 'SOUTH YORKSHIRE <<<')
} else {
  place_selection <- slopes.log1521 %>% filter(SIC07_description == sector) %>%
    arrange(slope)
    sy_position <<- length(place_selection$ITL_region_name) + 1 -which(place_selection$ITL_region_name == 'SOUTH YORKSHIRE <<<')
  }

#Keep only the top ten places and order them
timeplot <- timeplot %>% 
  mutate(ITL_region_name = factor(ITL_region_name, ordered = T, levels = place_selection$ITL_region_name)) %>%
  filter(ITL_region_name %in% c(place_selection$ITL_region_name[1:10],place))#Make sure chosen place added

#Mark the ITL of interest so it can be clearer in the plot
timeplot <- timeplot %>%
  mutate(
    ITL2ofinterest = ifelse(ITL_region_name == place, 'ITL of interest','other'),
  )

ggplot(timeplot %>% 
         rename(`ITL region` = ITL_region_name) %>% 
         filter(!is.na(movingav)),#remove NAs from dates so the x axis doesn't show them
       aes(x = year, y = value, colour = `ITL region`, size = ITL2ofinterest, linetype = ITL2ofinterest, group = `ITL region`)) +
       # aes(x = year, y = movingav, colour = `ITL region`, size = ITL2ofinterest, linetype = ITL2ofinterest, group = `ITL region`)) +
  geom_point() +
  geom_line() +
  scale_y_log10() +
  scale_size_manual(values = c(2.5,1)) +
  scale_color_brewer(palette = 'Paired', direction = 1) +
  ylab('Regional GVA chained volume (log, millions)') +
  guides(size = "none", linetype = "none") +
  ggtitle(
    paste0(ifelse(growing,'10 HIGHEST GROWTH PLACES','10 LOWEST GROWTH PLACES'),'\n',sector,'\n', place, ' highlighted (',scales::ordinal(sy_position),' most positive slope)')
  ) +
  theme(plot.title = element_text(face = 'bold'))
```


----

## Information and communication slope pair matrix 2015 to 2021

Slope pairs can be looked at for a single SIC section, comparing all pairs for every place in the UK. South Yorkshire is 8th from the top (and right) - for 2015-21, it has the clearest statistical separation (95% CIs again) from all other places. 

The strong green colours of the names on the axes (for all but Outer London) show ICT growing everywhere - but for this latest 7 year period, South Yorkshire has likely had significantly stronger growth.

The next slide shows the same plot for 2013-2019, where the same is not true.

```{r}
p <- slopeDiffGrid(slope_df = slopes.log1521, confidence_interval = 95, column_to_grid = ITL_region_name, column_to_filter = SIC07_description, filterval = 'Information and communication')

p + coord_fixed()

```

----

## Information and communication slope pair matrix 2013 to 2019

For the 7 years prior to COVID, everywhere saw ICT growth (strong green axis names) and there was a mix of places with separable growth (all growing, but some with growth slopes significantly shallower than other places). 

In this time period, South Yorkshire DOES NOT stand out (though ICT growth in SY is still strong, see 1st pair matrix plot). It's at least partly its continued strength through COVID in the following two years that change the story (though that 2018+ growth jump is clearly a factor).

```{r}
p <- slopeDiffGrid(slope_df = slopes.log1319, confidence_interval = 95, column_to_grid = ITL_region_name, column_to_filter = SIC07_description, filterval = 'Information and communication')

p + coord_fixed()

```

----







----

## Manufacturing time series: 10 HIGHEST growth places

Although South Yorkshire's manufacturing GVA growth pattern is similar to other places with strong growth, it does not really stand out statistically, as the next slides show. 

```{r, fig.width = 10, fig.height = 8}
growing = T
place = 'SOUTH YORKSHIRE <<<'
sector = 'Manufacturing'

timeplot <- itl2.cvs %>% 
      filter(SIC07_description == sector) 
    
#Use zoo's rollapply function to get a moving average
timeplot <- timeplot %>% 
  group_by(ITL_region_name) %>% 
  arrange(year) %>% 
  mutate(
    movingav = rollapply(value,3,mean,align='right',fill=NA)
  )

#Use largest positive log slopes to filter
if(growing){
  place_selection <- slopes.log1521 %>% filter(SIC07_description == sector) %>%
    arrange(-slope)
  sy_position <<- which(place_selection$ITL_region_name == 'SOUTH YORKSHIRE <<<')
} else {
  place_selection <- slopes.log1521 %>% filter(SIC07_description == sector) %>%
    arrange(slope)
    sy_position <<- length(place_selection$ITL_region_name) + 1 -which(place_selection$ITL_region_name == 'SOUTH YORKSHIRE <<<')
  }

#Keep only the top ten places and order them
timeplot <- timeplot %>% 
  mutate(ITL_region_name = factor(ITL_region_name, ordered = T, levels = place_selection$ITL_region_name)) %>%
  filter(ITL_region_name %in% c(place_selection$ITL_region_name[1:10],place))#Make sure chosen place added

#Mark the ITL of interest so it can be clearer in the plot
timeplot <- timeplot %>%
  mutate(
    ITL2ofinterest = ifelse(ITL_region_name == place, 'ITL of interest','other'),
  )

ggplot(timeplot %>% 
         rename(`ITL region` = ITL_region_name) %>% 
         filter(!is.na(movingav)),#remove NAs from dates so the x axis doesn't show them
       aes(x = year, y = value, colour = `ITL region`, size = ITL2ofinterest, linetype = ITL2ofinterest, group = `ITL region`)) +
       # aes(x = year, y = movingav, colour = `ITL region`, size = ITL2ofinterest, linetype = ITL2ofinterest, group = `ITL region`)) +
  geom_point() +
  geom_line() +
  scale_y_log10() +
  scale_size_manual(values = c(2.5,1)) +
  scale_color_brewer(palette = 'Paired', direction = 1) +
  ylab('Regional GVA chained volume (log, millions)') +
  guides(size = "none", linetype = "none") +
  ggtitle(
    paste0(ifelse(growing,'10 HIGHEST GROWTH PLACES','10 LOWEST GROWTH PLACES'),'\n',sector,'\n', place, ' highlighted (',scales::ordinal(sy_position),' most positive slope)')
  ) +
  theme(plot.title = element_text(face = 'bold'))
```


----

## Manufacturing slope pair matrix 2013 to 2019

While manufacturing in South Yorkshire has a significantly positive slope itself, it doesn't really stand out nationally. For the pre-COVID time period, compared to some places its growth slope is significantly lower, some others it's higher. 

```{r}
p <- slopeDiffGrid(slope_df = slopes.log1319, confidence_interval = 95, column_to_grid = ITL_region_name, column_to_filter = SIC07_description, filterval = 'Manufacturing')

p + coord_fixed()

```

----

## Manufacturing slope pair matrix 2015 to 2021

It's a similar story for manufacturing from 2015 to 2021, covering COVID. South Yorkshire has relatively grown compared to some of the worst-performing places, but that's true of a number of ITL2 regions - as can be seen by the fact that South Yorkshire shares its significant green squares with many other places in the vertical stripes.

```{r}
p <- slopeDiffGrid(slope_df = slopes.log1521, confidence_interval = 95, column_to_grid = ITL_region_name, column_to_filter = SIC07_description, filterval = 'Manufacturing')

p + coord_fixed()

```

----


# JOB COUNTS

----

## South Yorkshire job count slope grid 2015 to 21

Jobs data from BRES is less reliable due to being survey-based and rounded before release, but some statistical patterns do emerge from slope differences.

The South Yorkshire slope grid for job counts supports ICT as a growth sector, though transportation/storage also has clearly separable jobs growth trends (we know that's largely the doubling of warehousing jobs from 2015, from previous work).

Job counts for most public sectors (public admin / education / admin) have dropped relative to the largest job growth sectors, and relative also to health.


```{r}
jobslopes.log <- get_slope_and_se_safely(data = itl2.jobs %>% filter(DATE %in% 2015:2021), GEOGRAPHY_NAME,SIC_SECTION_NAME, y = log(COUNT), x = DATE)

p <- slopeDiffGrid(slope_df = jobslopes.log, confidence_interval = 95, column_to_grid = SIC_SECTION_NAME, column_to_filter = GEOGRAPHY_NAME, filterval = 'SOUTH YORKSHIRE <<<')

p + coord_fixed()
```

----

## ICT job slope matrix

Looking just at the ICT sector and job slope pairs, while South Yorkshire does have significant jobs growth slopes, those are shared with many other places whose growth contrasts with mostly common shrinking ITL2 zones like Berkshire, Cheshire etc.

South Yorkshire's uniqueness here appears to be more due to productivity increases in the last five years, rather than relatively high job increases (though job numbers are still going up).

```{r}
p <- slopeDiffGrid(slope_df = jobslopes.log, confidence_interval = 95, column_to_grid = GEOGRAPHY_NAME, column_to_filter = SIC_SECTION_NAME, filterval = 'Information and communication')

p + coord_fixed()

```

----

# GVA per job: trend matrices for key sectors

The first 2D plots comparing GVA and jobs on different axes were useful for an overview; the next two trend pair matrices look for any statistically separable patterns in GVA per job trends.

----

## GVA per job slope matrix for South Yorkshire 2017-21

Statistically separating out slopes for GVA per worker within South Yorkshire tells a similar story to above: ICT strongest, manufacturing some hint of statistically separable increases. 

One addition here is finance and education standing out - but note from the earlier 2D plot slides, those GVA per worker gains come from GVA and job counts dropping, just one slightly less than the other, so the productivity number goes up.

Electricity/gas is a tiny sector jobs and GVA wise; its growth is probably not relevant to SY's overall growth trajectory.

```{r}
perjobslopes.log <- get_slope_and_se_safely(data = itl2.gvaperjob %>% filter(year %in% 2017:2021), ITL_region_name,SIC07_description, y = log(gvaperjob), x = year)

#Functioning up
# debugonce(slopeDiffGrid)
slopeDiffGrid(slope_df = perjobslopes.log, confidence_interval = 95, column_to_grid = SIC07_description, column_to_filter = ITL_region_name, filterval = 'SOUTH YORKSHIRE <<<')
```

----

## GVA per job slope matrix for manufacturing 2017-2021

Manufacturing GVA per job slope differences for 2017-21 show how South Yorkshire (despite a significant growth slope) doesn't stand out from other places.

```{r}

slopeDiffGrid(slope_df = perjobslopes.log, confidence_interval = 95, column_to_grid = ITL_region_name, column_to_filter = SIC07_description, filterval = 'Manufacturing')

```

----

## GVA per job slope matrix for ICT 2017-2021

ICT, in comparison - both South Yorkshire and Greater Manchester stand out as having stronger GVA per worker growth than elsewhere.

```{r}

slopeDiffGrid(slope_df = perjobslopes.log, confidence_interval = 95, column_to_grid = ITL_region_name, column_to_filter = SIC07_description, filterval = 'Information and communication')

```









```{r CUTTINZ, eval = F}
# The raw numbers for South Yorkshire 2021 (ordered by GVA)

# These are the raw numbers, with highest GVA at the top, just 


job_gva_scale <- itl2.gvaperjob %>% 
  filter(year == 2021, ITL_region_name=='SOUTH YORKSHIRE <<<',SIC07_description!='Real estate activities') %>% 
  arrange(-gva) %>% 
  select(SIC07_description,gva,jobcount,gvaperjob)

ft <- flextable(job_gva_scale) %>%
  width(width = 1.5) %>%
  fontsize(size = 9) %>%
  fontsize(size = 9, part = 'header') %>%
  padding(padding = 0, part = "all") %>%
  autofit()

ft

# job_gva_scale

```


